---
title: "introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SARNr)
library(sf) #for vector processing
library(terra) #for raster processing
```

This vigentte will get you started extracting actively-flowing river networks from Planet imagery!

# Preprocessing
`SARNr` assumes you have the following data prepared:

- Multi-band remote sensing image with at least the following bands: green, red, and nir. We developed `SARNr` around Planet imagery and so these bands are the bare minimum, though other sensors can be used to calculate other spectral indices for river classification.
- DEM river network of equivalent (or better) resolution to your remotely sensed image(s). This resolution requirement is important to faciliate raster algebra!

One can also generate their DEM river network in a GIS: to my knowledge there is no easy way to do this in R currently. This needs to only be done once. The general workflow is the following (there are tons of tutorials online for doing this in ArcGIS, QGIS, GRASS, etc.)

- Fill sinks
- Generate flow directions
- Generate flow accumulation grid
- Set channelization threshold and create binary river network
- Convert this raster river network to vector

# Workflow
Below is an example code for generating an actively flowing drainage network given a DEM river network and a 3m Planet Labs satellite image from 10/09/2020 in the Nepalese Himalaya.

Load in the necessary data
```R
#load in data
rivnet_dem <- st_read('data/DEM_rivnet_33333_20201009_045732_0f17.shp') #DEM river networks
dem_domain <- rast('data/dem_fin_20201009_045732_0f17.tif') #dem downscaled to 3m
image <- rast('data/PSScene4Band-20201009_045732_0f17.tif') #3m PlanetScope image
```

Classify water within your image. You can do this using bespoke methods or one the function included here (see below). Here we use a bespoke combination of Rosgen unimodal thresholding (https://www.sciencedirect.com/science/article/pii/S0034425717301475) and a water classification method developed for Landsat in the Himalaya (https://doi.org/10.3390/s18082580).

If you are going to make your own river masks, simply make sure the output is a `terra::spatRaster` object with river pixels coded as 1s and land pixels as NaN. `SARNr` will throw an error if this format is not followed.

```R
#Rosgen's Unimodal thresholding approach applied to the spectral index (NDWI-NDVI)
rivnet_rs_rast <- sarn_classifyWater_unimodal(img = image,
                                     dem = dem_domain,
                                     maxElev = 4000) #water is ignored above this elevation (to remove snow/ice/glaciers)
```

Now that you have a water mask, run the algorithm
```R
#run SARNr
data <- sarn_data(dem_network = rivnet_dem,
                  riverMask = rivnet_rs_rast,
                  dem = dem_domain) #create SARN object

joinedNetwork <- sarn_joinNetworks(data = data, 
                                   bufferSize=5) #join DEM and RS networks

trimmedNetwork <- sarn_trimNetworks(networks = joinedNetwork,
                                    printOutput = 'Yes') #trim combined DEM/RS network

fin <- sarn_hydrography(trimmedNetwork = trimmedNetwork$trimmedNetwork, 
                        dem = data$dem,
                        riverMask = data$riverMask, 
                        lengthThresh = 50,
                        printOutput = 'Yes') #build hydrography

#save river network to shp
st_write(fin, 'debug/hydrography.shp', delete_dsn = TRUE)
```
